\textbf{ Mapping stellar surfaces\\ I: Degeneracies in the rotational light curve problem }

### 2021-01-28 paper1/ms.tex
adding some words about information and cramer-rao

\subsection{Implications for inference}
%
The procedure we carried out in the previous section is similar to what was done in \citet{Rauscher2018}
for eclipse mapping of exoplanets. That work decomposed the surface map
using principal component analysis (PCA), finding an orthogonal set of
light curve signals and their corresponding surface modes (the left and right
singular vectors in SVD, respectively). Inference was then carried out in
the basis spanned by the components in the preimage, which can be uniquely
constrained by the data.

Let us therefore consider Figure~\ref{fig:nullspace_preimage} in the
context of an inference exercise. One can equivalently think of
the preimage as the solution to a least squares inference problem
in which a completely uninformative prior%
\footnote{%
    More specifically, a zero-mean, infinite variance Gaussian prior on the coefficients of $\mathbf{y}$.
}
is placed on the spherical harmonic
coefficients. In this case, the least squares solution is given by
%
\begin{proof}{test_svd_lstsq}
    \label{eq:inference}
    \hat{\mathbf{y}}
    &=
    \lim_{\lambda \rightarrow 0} \left( \pmb{\mathcal{A}}^\top \pmb{\mathcal{A}} + \lambda \mathbf{I}\right)^{-1} \pmb{\mathcal{A}}^\top (\mathbf{f} - \mathbf{1})
    \nonumber\\
    &= \mathbf{y}_\bullet
    \quad.
\end{proof}
%
In this sense, $\mathbf{y}_\bullet$ represents our knowledge about the
surface of a star after an observation if we have no prior information
whatsoever on $\mathbf{y}$.
\emph{Any other information about the surface must come from the prior.}%

@@

The effect of different priors on the structure and size of the null
space is beyond the scope of this paper. In general, this is a particularly
difficult question to address because non-Gaussian priors on the
surface map break the linearity of the problem. While compactness priors
(like the assumption of a small number of discrete circular spots) can
break many of the degeneracies discussed here, simulation-based
arguments show that it is still not possible to uniquely constrain
their number, locations, or sizes from individual light curves
\citep{Basri2020}.

>>>>

\subsection{Implications for inference}
%
The procedure we carried out in the previous section is similar to what was done in \citet{Rauscher2018}
for eclipse mapping of exoplanets. That work decomposed the surface map
using principal component analysis (PCA), finding an orthogonal set of
light curve signals and their corresponding surface modes (the left and right
singular vectors in SVD, respectively). Inference was then carried out in
the basis spanned by the components in the preimage, which can be uniquely
constrained by the data.
\dfm{I think this paragraph needs a little more... e.g. This is all good and well when the inclination, rotation period, and limb darkening are known, so the data can be uniquely projected. But that is no longer true when these are parameters. Or is there more to it?}

The model mapping spherical harmonic coefficients to observed fluxes described in the preceding sections is linear, so we might expect that we could fit for the spherical harmonics using linear least-squares
%
\begin{align}
    \label{eq:linlstsq}
    \hat{\mathbf{y}} & \stackrel{?}{=} \left(\pmb{\mathcal{A}}^\top\,\pmb{\mathcal{A}}\right)^{-1}\,\pmb{\mathcal{A}}^\top\,\left(\mathbf{f} - \mathbf{1}\right)
    \quad.
\end{align}
%
But, since the flux operator $\pmb{\mathcal{A}}$ is low-rank, the model is underdetermined, so the above operation is not defined and, equivalently, the Fisher information matrix
%
\begin{align}
    \label{eq:fisher-info}
    \pmb{\mathcal{I}}(\mathbf{y}) & = \frac{1}{\sigma^2} \, \pmb{\mathcal{A}}^\top\,\pmb{\mathcal{A}}
\end{align}
%
is singular.
In Equation~(\ref{eq:fisher-info}), $\sigma$ is the flux measurement uncertainty.
There exist many methods and procedures for linear regression for underdetermined models, but all solutions amount to imposing stronger assumptions about the model in order to break the degeneracies.

One standard method is regularized least squares where Equation~(\ref{eq:linlstsq}) becomes
%
\begin{align}
    \hat{\mathbf{y}}_\lambda & \equiv \left(\pmb{\mathcal{A}}^\top\,\pmb{\mathcal{A}} + \lambda\,\mathbf{I}\right)^{-1}\,\pmb{\mathcal{A}}^\top\,\left(\mathbf{f} - \mathbf{1}\right)
\end{align}
%
where $\mathbf{I}$ is the identity matrix and $\lambda$ is a parameter controlling the strength of the regularization.
In a Bayesian context, this can be interpreted as placing a Gaussian prior with inverse variance $\sigma^2\,\lambda$ on the spherical harmonic coefficients.

In the limit of infinitesimal regularization, it can be demonstrated \citep[see][for example]{Hogg2021} \dfm{(TODO: I think I know how to prove this)} that
%
\begin{proof}{test_svd_lstsq}
    \label{eq:inference}
    \lim_{\lambda \rightarrow 0_+} \hat{\mathbf{y}}_\lambda
    % &=
    % \lim_{\lambda \rightarrow 0} \left( \pmb{\mathcal{A}}^\top \pmb{\mathcal{A}} + \lambda \mathbf{I}\right)^{-1} \pmb{\mathcal{A}}^\top (\mathbf{f} - \mathbf{1})
    % \nonumber\\
    &= \mathbf{y}_\bullet
    \quad.
\end{proof}
%
In this sense, $\mathbf{y}_\bullet$ represents our knowledge about the
surface of a star after an observation if we have no prior information
whatsoever on $\mathbf{y}$.
\emph{Any other information about the surface is entirely driven by our assumptions.}%

It is interesting to quantify the information content of a light curve as a function of spherical harmonic coefficients.
While Equation~(\ref{eq:fisher-info}) provides a compact expression for the Fisher information matrix, the most important feature that we want to quantify is the covariance between coefficients.
This is most naturally captured by the covariance matrix of the likelihood function for the spherical harmonic coefficients:
%
\begin{align}
    \pmb{\Sigma}_\lambda & = \sigma^2\,\left(\pmb{\mathcal{A}}^\top\,\pmb{\mathcal{A}} + \lambda\,\mathbf{I}\right)^{-1} \quad.
\end{align}
%
Below, we use this to quantify the width of the likelihood for each spherical harmonic coefficient by plotting the ``relative variance'' $S$, which we define as the diagonal of the matrix $\pmb{\Sigma}_\lambda$ in units of $\lambda / \sigma^2$ in the limit where $\lambda \to 0_+$.
This choice of units constrains $0 \le S \le 1$ and allows easy comparisons between the information content of different datasets without being sensitive to the specific modeling assumptions.
Under this definition, when $S$ approaches one for a particular coefficient, this means that the likelihood function provides no information about that coefficient, while $S = 0$ indicates that that coefficient can be perfectly measured.

This relative variance $S$ can be interpreted either as the Cram\'er--Rao lower bound on the variance of an estimator for a spherical harmonic \citep[e.g.,][]{Li2012}, although our specific choice of units are somewhat non-standard.
Similarly, in a Bayesian context, this is the ``posterior shrinkage'', the ratio of the posterior variance to the prior variance in the limit of infinite prior variance \citep[e.g.,][]{Betancourt2018}.

@@

The effect of different modeling assumptions on the structure and size of the null
space is beyond the scope of this paper. In general, this is a particularly
difficult question to address because non-Gaussian priors on the
surface map break the linearity of the problem. While compactness constraints
(like the assumption of a small number of discrete circular spots) can
break many of the degeneracies discussed here, simulation-based
arguments show that it is still not possible to uniquely constrain
their number, locations, or sizes from individual light curves
\citep{Basri2020}.
